---
title: "My Project"
author: "Can Zhang"
date: "2024-10-10"
output: "github_document"
---
# Load Packages and Dataset
```{r}

library(haven)
library(dplyr)
library(ggplot2)
library(tidyr)
library(car)
library(moments)
library(boot)
load("C:/Users/User/Downloads/ICPSR_38964-V2/ICPSR_38964/DS0001/38964-0001-Data.rda")
dataset <- da38964.0001
rm(da38964.0001)

```

# Data Cleaning
```{r}

#Select variables related to my research question
my_dataset <- dataset %>% select(RELIG_ATTEND, RELIG_IMP, RACEREC, LONELY_A, LONELY_B, LONELY_C)
my_dataset <- drop_na(my_dataset)
summary(my_dataset)

#recode and composite the variables
my_dataset$RELIG_ATTEND <- 5 - as.numeric(my_dataset$RELIG_ATTEND)
my_dataset$RELIG_IMP <- 5 - as.numeric(my_dataset$RELIG_IMP)
my_dataset$RELIG_ENGAGEMENT <- (my_dataset$RELIG_ATTEND + my_dataset$RELIG_IMP) / 2

my_dataset$LONELY_A <- as.numeric(my_dataset$LONELY_A)
my_dataset$LONELY_B <- as.numeric(my_dataset$LONELY_B)
my_dataset$LONELY_C <- as.numeric(my_dataset$LONELY_C)
my_dataset$LONELY <- (my_dataset$LONELY_A + my_dataset$LONELY_B + my_dataset$LONELY_C) / 3

```

# Statistical Assumptions and Data Transformation
# Normality Assumption
```{r}

# Check the dependent variable loneliness
# Density plot of loneliness
ggplot(my_dataset, aes(x = LONELY)) + geom_density(adjust = 1.5) + facet_wrap(~RACEREC) + theme_classic()

# Check the independent variable religious engagement
# Skewness and kurtosis for religious engagement
skewness(my_dataset$RELIG_ENGAGEMENT)
kurtosis(my_dataset$RELIG_ENGAGEMENT)

```

The distribution of loneliness is positively skewed, as indicated by the density plot showing a positive skew.

The distribution of religious engagement is slightly less concentrated around the mean and has fewer extreme values in the tails compared to a normal distribution, as the kurtosis is lower than 3 (1.83). However, since the sample size is large enough, it is generally acceptable to have such kurtosis.

# Equal Variance Assumption
```{r}

leveneTest(LONELY ~ RACEREC, data = my_dataset)

```

The Levene's Test result is not significant, indicating that the variance is equal across groups (assumption is met).

# Data Transformation
```{r}

my_dataset$LONELY_log <- log(my_dataset$LONELY)

ggplot(my_dataset, aes(x = LONELY_log)) + geom_density(adjust = 1.5) + facet_wrap(~RACEREC) + theme_classic()

```

After the log transformation, the data distribution of loneliness is generally normal, according to the density plot.