---
title: "My Project"
author: "Can Zhang"
date: "2024-10-24"
output: "github_document"
---
# Load Packages and Dataset
```{r}

library(haven)
library(dplyr)
library(ggplot2)
library(tidyr)
library(car)
library(moments)
library(boot)
library(bruceR)
library(interactions)
library(ggrepel)
load("C:/Users/zhang/OneDrive/2024-25/PSY329/ICPSR_38964-V2/ICPSR_38964/DS0001/38964-0001-Data.rda")
dataset <- da38964.0001
rm(da38964.0001)

```

# Data Cleaning
```{r}

#Select variables related to my research question
my_dataset <- dataset %>% select(RELIG_ATTEND, RELIG_IMP, RACEREC, LONELY_A, LONELY_B, LONELY_C)
my_dataset <- drop_na(my_dataset)
summary(my_dataset)

#recode and composite the variables
my_dataset$RELIG_ATTEND <- 6 - as.numeric(my_dataset$RELIG_ATTEND)
my_dataset$RELIG_IMP <- 5 - as.numeric(my_dataset$RELIG_IMP)
my_dataset$RELIG_ENGAGEMENT <- (my_dataset$RELIG_ATTEND + my_dataset$RELIG_IMP) / 2

my_dataset$RACEREC <- gsub("Non-Hispanic ", "", my_dataset$RACEREC)

my_dataset$LONELY_A <- as.numeric(my_dataset$LONELY_A)
my_dataset$LONELY_B <- as.numeric(my_dataset$LONELY_B)
my_dataset$LONELY_C <- as.numeric(my_dataset$LONELY_C)
my_dataset$LONELY <- (my_dataset$LONELY_A + my_dataset$LONELY_B + my_dataset$LONELY_C) / 3

race_summary <- my_dataset %>%
  group_by(RACEREC) %>%
  summarize(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) 

ggplot(race_summary, aes(x = "", y = percentage, fill = RACEREC)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_label_repel(aes(label = paste0(round(percentage, 1), "%")),
                   position = position_stack(vjust = 0.5),
                   box.padding = 0.5,
                   show.legend = FALSE) +
  labs(title = "Race Composition of Participants", fill = "Race/Ethnicity") +
  theme_minimal(base_size = 14) +
  theme(axis.text = element_blank(), axis.title = element_blank(), panel.grid = element_blank()) +
  theme(plot.background = element_rect(fill = "#f6c1ad", color = NA)) + 
  scale_fill_brewer(palette = "Set3")

```

# Statistical Assumptions and Data Transformation
# Normality Assumption
```{r}

# Check the dependent variable loneliness
# Density plot of loneliness
ggplot(my_dataset, aes(x = LONELY)) + geom_density(adjust = 1.5) + facet_wrap(~RACEREC) + theme_classic()

# Check the independent variable religious engagement
# Skewness and kurtosis for religious engagement
skewness(my_dataset$RELIG_ENGAGEMENT)
kurtosis(my_dataset$RELIG_ENGAGEMENT)

```

The distribution of loneliness is positively skewed, as indicated by the density plot showing a positive skew.

The distribution of religious engagement is slightly less concentrated around the mean and has fewer extreme values in the tails compared to a normal distribution, as the kurtosis is lower than 3 (1.83). However, since the sample size is large enough, it is generally acceptable to have such kurtosis.

# Equal Variance Assumption
```{r}

leveneTest(LONELY ~ RACEREC, data = my_dataset)

```

The Levene's Test result is not significant, indicating that the variance is equal across groups (assumption is met).

# Data Transformation
```{r}

my_dataset$LONELY_log <- log(my_dataset$LONELY)

ggplot(my_dataset, aes(x = LONELY_log)) + geom_density(adjust = 1.5) + facet_wrap(~RACEREC) + theme_classic()

```

After the log transformation, the data distribution of loneliness is generally normal, according to the density plot.

# Data Analysis
```{r}

cor_data <- my_dataset %>% select(RELIG_ENGAGEMENT, LONELY_log, RACEREC)
Corr(cor_data)

PROCESS(my_dataset, y = "LONELY_log", x = "RELIG_ENGAGEMENT", mods = c("RACEREC"))

```

# Visualize the Results
```{r}

ggplot(cor_data, aes(x = LONELY_log, y = RELIG_ENGAGEMENT)) + geom_point(alpha = .3) + geom_smooth(method = "lm", se = TRUE, linetype = "solid", color = "#e7553d") +  theme_minimal(base_size = 14)

ggplot(data = cor_data, aes(x = LONELY_log, y = RELIG_ENGAGEMENT)) +
  geom_point(alpha = .3) +
  geom_smooth(method = "lm", se = TRUE, linetype = "solid", color = "#e7553d") +
  facet_wrap(~RACEREC) +
  labs(title = "Correlation between Religious Engagement and Loneliness",
       subtitle = "Faceted by Race/Ethnicity",
       x = "Log-transformed Loneliness", 
       y = "Religious Engagement") +
  theme_minimal(base_size = 14)

```

# Reliability and Factor Analysis
```{r}

# Cronbach's Alpha
Alpha(my_dataset, vars = c("LONELY_A", "LONELY_B", "LONELY_C"))

# Factor Analysis
EFA(my_dataset, vars = c("LONELY_A", "LONELY_B", "LONELY_C"), method = "pa", plot.scree = TRUE, nfactors = c("parallel"))

# Spearman's Rho
spearman_rho <- cor(my_dataset$RELIG_ATTEND, my_dataset$RELIG_IMP, method = "spearman")
print(spearman_rho)

```

# Exploratory Questions
How do education affect the relationship between religious engagement and loneliness?
```{r}

my_dataset$EDUC <- dataset$EDUC
summary(my_dataset$EDUC)

leveneTest(LONELY_log ~ EDUC, data = my_dataset)
my_dataset$LONELY_sqrt <- sqrt(my_dataset$LONELY)
leveneTest(LONELY_sqrt ~ EDUC, data = my_dataset)
skewness(my_dataset$LONELY_sqrt)
kurtosis(my_dataset$LONELY_sqrt)

cor_data1 <- my_dataset %>% select(RELIG_ENGAGEMENT, LONELY_sqrt, EDUC)
Corr(cor_data1)

PROCESS(my_dataset, y = "LONELY_sqrt", x = "RELIG_ENGAGEMENT", mods = c("EDUC"))

ggplot(data = my_dataset, aes(x = RELIG_ENGAGEMENT, y = LONELY_sqrt, color = EDUC)) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~ EDUC, ncol = 3, scales = "fixed") + 
  labs(title = "Effect of Religious Engagement on Loneliness by Education",
       x = "Religious Engagement",
       y = "Sqrt-transformed Loneliness") +
  theme_minimal(base_size = 14) + theme(strip.text = element_text(size = 10)) + 
  theme(plot.background = element_rect(fill = "#FDFDF3", color = NA)) + 
  theme(legend.position = "none")

```

```{r}

model_main <- lm(LONELY_log ~ RELIG_ENGAGEMENT + RACEREC, data = my_dataset)
summary(model_main)

model_interaction <- lm(LONELY_log ~ RELIG_ENGAGEMENT * RACEREC, data = my_dataset)
summary(model_interaction)

model_control <- lm(LONELY_log ~ RELIG_ENGAGEMENT * RACEREC + EDUC, data = my_dataset)
summary(model_control)

# Residual plot for homoscedasticity
plot(model_control, which = 1)

# Q-Q plot for normality of residuals
plot(model_control, which = 2)

interact_plot(model_control, pred = RELIG_ENGAGEMENT, modx = RACEREC, interval = TRUE)

filtered_data <- cor_data %>%
  filter(RACEREC != "(98) Don't know")
ggplot(data = filtered_data, aes(x = RELIG_ENGAGEMENT, y = LONELY_log, color = RACEREC)) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~ RACEREC, ncol = 3, scales = "fixed") + 
  labs(title = "Effect of Religious Engagement on Loneliness by Race",
       x = "Religious Engagement",
       y = "Log-transformed Loneliness") +
  theme_minimal(base_size = 14) + theme(strip.text = element_text(size = 10)) + 
  theme(plot.background = element_rect(fill = "#FDFDF3", color = NA), legend.position = "none")
```